---
title: "Loops and Functions"
author: "Somraj Shrestha"
date: "2026-03-26"
output: 
  md_document:
    variant: gfm

---

1.	1 pts. Regarding reproducibility, what is the main point of writing your own functions and iterations? 
- make the analysis consistent, repeatable, and less error-prone.

2.	2 pts. Conceptual: In your own words, describe: 1) how to write a function in R, and 2) how to write a for loop in R. In your answer, be specific about: basic syntax, where the code is written, and how results are returned or stored. (Answer in plain text; no code required for this question.)
- In R,while writing a function,I do function_name<-function(input variables){code} i.e I start by giving the function a name and use the function() command to define it. I put any inputs the function needs inside the parentheses, and I write the code it should run inside curly braces { }. If I want the function to give me back a result, I use return() at the end.

- The way to write a for loop in R is :  for (variable in sequence) { code }. Inside the parentheses, I choose a variable name (usually "i") that will represent each item in the sequence, and I define the range or list I want to loop through. For a range, I use something like 1:10 to go from 1 to 10. For a list or vector, I can use something like c("A", "B", "C") to loop through each item. I can also use base R functions like rep(), seq() or seq_along() as per need to define the range or list. The code inside the curly braces { } is what I want to run during each loop. Each time the loop runs, the variable takes on the next value in the sequence or list, and the code runs using that value.

3.	1 pts. Read in the Cities.csv file from Canvas using a relative file path.
```{r}
citydata <- read.csv("Cities.csv", na.strings = "na")
```

4.	6 pts. Function writing: Write a function that calculates the distance (in kilometers) between two pairs of geographic coordinates using the Haversine formula. 
a.	Function requirements:
i.	Inputs: lat1, lon1, lat2, lon2
ii.	The function must return a single numeric value called distance_km
iii.	All code shown below must be placed inside the function body


```{r Creating function}
dist<- function(lat1, lon1, lat2, lon2) {
  # convert to radians
  rad.lat1 <- lat1 * pi / 180
  rad.lon1 <- lon1 * pi / 180
  rad.lat2 <- lat2 * pi / 180
  rad.lon2 <- lon2 * pi / 180

  # Haversine formula
  delta_lat <- rad.lat2 - rad.lat1
  delta_lon <- rad.lon2 - rad.lon1
  a <- sin(delta_lat / 2)^2 + cos(rad.lat1) * cos(rad.lat2) * sin(delta_lon / 2)^2
  c <- 2 * asin(sqrt(a)) 

  # Earth's radius in meters
  earth_radius <- 6378137

  # Calculate the distance in kilometers
  distance_km <- (earth_radius * c) / 1000

  return(distance_km)
}
```

5.	6 pts. Using your function from question 4, calculate the distance between Auburn, AL and New York City.
a.	Subset or extract only the latitude and longitude values needed from Cities.csv
b.	Pass those values into your function
c.	Store the result in an object

The expected output is 1367.854 km (small rounding differences are acceptable).


```{r}
library(tidyverse)
distance <- citydata %>%
  filter(city %in% c("Auburn", "New York")) %>%
  summarise(distance_km = dist(
    lat[city == "Auburn"],
    long[city == "Auburn"],
    lat[city == "New York"],
    long[city == "New York"]
  ))
distance
```

6.	4 pts. Use your distance function inside a for loop to calculate the distance between Auburn, AL and every other city in Cities.csv. Each iteration should calculate the distance for one city
```{r}
library(tidyverse)

for (i in seq_along(citydata$city)) { # Here I used seq_along to loop from 1st to 40th city
      lat1 <- citydata$lat[citydata$city=="Auburn"]
      lon1 <- citydata$long[citydata$city=="Auburn"]
      lat2 <- citydata$lat[i]#Assigning latitudes of each city
      lon2 <- citydata$long[i]#Assigning longitude of each city

      distance_km <- dist(lat1, lon1, lat2, lon2) # using the dist function I created to calculate distance
      print(distance_km)
}


```

7.	4 pts. Modify your loop so that each iteration appends one new row to a dataframe with the following columns:

•	City1 (the other city)
•	City2 (Auburn)
•	Distance_km

```{r}
library(tidyverse)

distance_df2 <- NULL #Creating a new data frame 

for (i in seq_along(citydata$city)) { # Here I used seq_along to loop from 1st to 40th city
      lat1 <- citydata$lat[citydata$city=="Auburn"]
      lon1 <- citydata$long[citydata$city=="Auburn"]
      lat2 <- citydata$lat[i]
      lon2 <- citydata$long[i]

      distance_km <- dist(lat1, lon1, lat2, lon2)

      distance_df2 <- rbind( # Appending each city combination and their distances into the data frame after each iteration
        distance_df2,
        data.frame(
          city1 = "Auburn",
          city2 = citydata$city[i],
          distance_km = distance_km
        ))
}


distance_df2
```



8.	1 pts. Commit and push a gfm .md file to GitHub inside a directory called Coding Challenge 6. Provide me a link to your github written as a clickable link in your rendered .pdf, .docx, or .html

[Link to my github]

