---
title: "DataVisualization_Challenge1"
author: "Somraj Shrestha"
date: "`r Sys.Date()`"
output: html_document
---

1.	5 pts. Explain the following concepts about ggplot.

a. What three elements do you need to produce a ggplot?
- dataframe (data), aesthetic mapping (x and y variables, shape, color etc.), and a geom (the type of plot).

b.	What is a geom? 
- It is a geometric object which determines how the data are drawn on the plot.

c.	What is a facet? 
- It helps spliting the plots into small panels based on categorical variable thus, allowing us to compare between the groups within a dataframe.

d.	Explain the concept of layering.
- Layering means building a plot layer by layer by adding components that we need for the plot using +. Each layer adds information to the plot.

e.	Where do you add x and y variables and map different shapes, colors, and other attributes to the data? 
- We add x and y variables and different shapes, colors, and attributes inside aes() in the ggplot(). Or we can add them inside a specific geom too.

2.	4 pts. Make a boxplot using ggplot with DON as the y variable, treatment as the x variable, and color mapped to the wheat cultivar. Change the y label to “DON (ppm)” and make the x label blank.

```{r}
myco_data <- read.csv("Data/MycotoxinData.csv", na.strings = "na") #load the data
str(myco_data)
```
```{r}
myco_data$Treatment <- as.factor(myco_data$Treatment)
myco_data$Cultivar <- as.factor(myco_data$Cultivar)
str(myco_data)
```


```{r}
library(ggplot2)
plot1 <- ggplot(myco_data, aes(x = Treatment,  y = DON, colour = Cultivar)) + 
  geom_boxplot() +       #making a boxplot
  xlab("") +             #labelling X-axis
  ylab("DON(ppm)")       #labelling Y-axis
plot1
```

3.	4 pts. Now convert this data into a bar chart with standard-error error bars using the stat_summary() command. (hint: use position = dodge)

```{r}
plot2 <- ggplot(myco_data, aes(x = Treatment,  y = DON, colour = Cultivar)) + 
   stat_summary(fun = mean, geom = "bar", position = "dodge") +  #genrating barchart with mean
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge")  #adding standard error bars on the bars
plot2
```
4.	3 pts. Add points to the foreground of the boxplot and bar chart you made in questions 2 & 3 that show the distribution of points. Set the shape = 21 and the outline color black (hint: use jitter_dodge). 

```{r}
plot3 <- ggplot(myco_data, aes(x = Treatment,  y = DON, colour = Cultivar)) + 
  geom_boxplot() +
  geom_point(position = position_jitterdodge(), shape = 21, color = "black") + #adding points on the box-plot to visualize the data better
  xlab("") +
  ylab("DON(ppm)")
plot3
```

```{r}
plot4 <- ggplot(myco_data, aes(x = Treatment,  y = DON, colour = Cultivar)) +
   stat_summary(fun = mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
   geom_point(position = position_jitterdodge(), shape = 21, color = "black")
plot4
```

5.	2 pts. Add a facet to the plots based on cultivar.

```{r}
plot5 <- ggplot(myco_data, aes(x = Treatment,  y = DON, colour = Cultivar)) + 
  geom_boxplot() +
  geom_point(position = position_jitterdodge(), shape = 21, color = "black") +
  xlab("") +
  ylab("DON(ppm)") +
  facet_wrap(~Cultivar, scales = "free") #adding facet to divide the dataframe into sub-groups
plot5
```
```{r}
plot6 <- ggplot(myco_data, aes(x = Treatment,  y = DON, colour = Cultivar)) +
   stat_summary(fun = mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
   geom_point(position = position_jitterdodge(), shape = 21, color = "black") +
  facet_wrap(~Cultivar, scales = "free")
plot6
```

6.	3 pts. Add transparency to the points so you can still see the boxplot or bar in the background. 
```{r}
plot7 <- ggplot(myco_data, aes(x = Treatment,  y = DON, colour = Cultivar)) + 
  geom_boxplot() +
  geom_point(position = position_jitterdodge(), shape = 21, color = "black") + 
  geom_jitter(width = 0.5, alpha = 0.5) + #adding transparency to the points
xlab("") +
  ylab("DON(ppm)") +
  facet_wrap(~Cultivar, scales = "free")
plot7
```
```{r}
plot8 <- ggplot(myco_data, aes(x = Treatment,  y = DON, colour = Cultivar)) +
   stat_summary(fun = mean, geom = "bar", position = "dodge") +
  stat_summary(fun.data = mean_se, geom = "errorbar", position = "dodge") +
   geom_point(position = position_jitterdodge(), shape = 21, color = "black") +
   geom_jitter(width = 0.5, alpha = 0.5) + #adding transparency to the points
  facet_wrap(~Cultivar, scales = "free")
plot8
```


7.	4 pts. Annotate your code and push it to github. 
Code annotated.
