---
title: "LinearModels:Homework"
author: "Somraj Shrestha"
date: "2026-03-26"
output: pdf_document
---

```{r}
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
```


```{r}
data("mtcars")
```

##Continuos variables
```{r}
ggplot(mtcars, aes(x = wt, y = mpg)) +
  geom_smooth(method = lm, se = FALSE, color = "grey") +
  geom_point(aes(color = wt)) +
  xlab("Weight") + 
  ylab("Miles per gallon") +
  scale_colour_gradient(low = "forestgreen", high = "black") +
  theme_classic()
```

There appears to be a linear relationship between wt and mpg. 

```{r}
d1 <- lm(mpg~wt, data = mtcars)
summary(d1)
```
In this case, about 75% variation in y is explained by x.

```{r}
anova(lm(mpg~wt, data = mtcars))
```
Since p = 1.29e-10, there is a significant relationship between our x and y variables.

```{r}
cor.test(mtcars$wt, mtcars$mpg)
```
```{r}
model <- lm(mpg~wt, data = mtcars)

ggplot(model, aes(y = .resid, x = .fitted)) +
  geom_point() +
  geom_hline(yintercept = 0)
```

##Categorical variables
```{r}
bull.rich <- read.csv("Data/Bull_richness.csv")
```

```{r}
bull.rich %>%
  filter(GrowthStage == "V8" & Treatment == "Conv.") %>%
  ggplot(aes(x = Fungicide, y = richness)) + 
  geom_boxplot()
```
```{r}
bull.rich.sub <- bull.rich %>%
  filter(GrowthStage == "V8" & Treatment == "Conv.")

t.test(richness~Fungicide, data = bull.rich.sub)
```
```{r}
t.test(richness~Fungicide, data = bull.rich.sub, var.equal = TRUE)
```

```{r}
summary(lm(richness~Fungicide, data = bull.rich.sub))
```
```{r}
anova(lm(richness~Fungicide, data = bull.rich.sub))
```
Assuming equal variance in groups and performing a two-sample t-test is the same result as a linear model and ANOVA on the linear model. The p-value tells us that the slope of the line is different than 0.

##ANOVAS- Continuous Y and multiple Categorical X.

```{r}
bull.rich.sub2 <- bull.rich %>%
  filter(Fungicide == "C" & Treatment == "Conv." & Crop == "Corn")
```

```{r}
bull.rich.sub2$GrowthStage <- factor(bull.rich.sub2$GrowthStage, levels = c("V6", "V8", "V15"))

ggplot(bull.rich.sub2, aes(x = GrowthStage, y = richness)) +
  geom_boxplot()
```

```{r}
lm.growth <- lm(richness ~ GrowthStage, data = bull.rich.sub2)
summary(lm.growth)
```

```{r}
anova(lm.growth)
```

```{r}
summary(aov(richness ~ GrowthStage, data = bull.rich.sub2))
```

```{r}
lsmeans <- emmeans(lm.growth, ~GrowthStage) # estimate lsmeans of variety within siteXyear
Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE) # contrast with Tukey ajustment by default. 
Results_lsmeans
```

##Understanding Interactions
```{r}
bull.rich.sub3 <- bull.rich %>%
  filter(Treatment == "Conv." & Crop == "Corn")

bull.rich.sub3$GrowthStage <- factor(bull.rich.sub3$GrowthStage, levels = c("V6", "V8", "V15"))
```

```{r}
# write it like this
lm.inter <- lm(richness ~ GrowthStage + Fungicide + GrowthStage:Fungicide, data = bull.rich.sub3)

# or like this
lm(richness ~ GrowthStage*Fungicide, data = bull.rich.sub3)
```

```{r}
summary(lm.inter) # significant terms
```

```{r}
anova(lm.inter) # The interaction term is signifant.
```

```{r}
bull.rich.sub3 %>%
  ggplot(aes(x = GrowthStage, y = richness, fill = Fungicide)) +
  geom_boxplot()
```

```{r}
lsmeans <- emmeans(lm.inter, ~Fungicide|GrowthStage) # estimate lsmeans of variety within siteXyear
Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE) # contrast with Tukey ajustment
Results_lsmeans
```


##Mixed Effect models
```{r}
lme0 <- lm(richness ~ GrowthStage*Fungicide, data = bull.rich.sub3)
lme1 <- lmer(richness ~ GrowthStage*Fungicide + (1|Rep), data = bull.rich.sub3)
summary(lme1)
```

```{r}
summary(lme0)
```

```{r}
summary(lme1)
```

