---
title: "CodingChallenge7"
author: "Somraj Shrestha"
date: "2026-03-26"
output:
  html_document:
    toc: true
    df_print: paged
  pdf_document:
    toc: true
---

1.	4 pts. Read in the data called “PlantEmergence.csv” using a relative file path and load the following libraries. tidyverse, lme4, emmeans, multcomp, and multcompView. Turn the Treatment , DaysAfterPlanting  and Rep into factors using the function as.factor
```{r}
datum1 <- read.csv("PlantEmergence.csv", na.strings = "na")
str(datum1)
```
```{r}
library(tidyverse)
library(lme4)
library(emmeans)
library(multcomp)
library(multcompView)
```

```{r}
datum1$Rep <- as.factor(datum1$Rep)
datum1$DaysAfterPlanting <- as.factor(datum1$DaysAfterPlanting)
datum1$Treatment <- as.factor(datum1$Treatment)
str(datum1)
```

2.	5 pts. Fit a linear model to predict Emergence using Treatment and DaysAfterPlanting along with the interaction. Provide the summary of the linear model and ANOVA results. 
```{r}
result1 <- lm(Emergence ~ Treatment + DaysAfterPlanting + Treatment:DaysAfterPlanting, data = datum1)
summary(result1)
```
```{r}
anova(result1)
```


3.	5 pts. Based on the results of the linear model in question 2, do you need to fit the interaction term? Provide a simplified linear model without the interaction term but still testing both main effects. Provide the summary and ANOVA results. Then, interpret the intercept and the coefficient for Treatment 2.

No, we don't need to fit the interaction term because the result is not significant (pvalue is equal to 1) meaning we can leave out the interaction term.

```{r}
result2 <- lm(Emergence ~ Treatment + DaysAfterPlanting, data = datum1)
summary(result2)
```

```{r}
anova(result2)
confint(result2)
```
We found that Treatment 2 had 134.53 (+-6.775364;+-95% CI) less emergence than Treatment 1 (pvalue< 2e-16).
The intercept is the average emergence for treatment 1 7 days after planting which is 182.163.
The coefficient of Treatment 2 is the difference in emergence between treatment 2 and treatment 1.

4.	5 pts. Calculate the least square means for Treatment using the emmeans package and perform a Tukey separation with the compact letter display using the cld function. Interpret the results.
```{r}
lsmeans <- emmeans(result2, ~Treatment) 
results_lsmeans <- cld(lsmeans, alpha = 0.05, details = TRUE) 
results_lsmeans
```

Treatment 2 (mean = 55.6) had significantly lower emergence than all others. Treatment 7 (mean = 188.0) had higher emergence than Treatment 2 but its emergence was statistically similar to Treatments 1, 9, 4, and 8. The highest emergence was in Treatments 5  which is statistically similar wtih Treatment 6, 3, 8, 4, 9, and 1.   While Treatments 6, 3, and 5 outperformed Treatment 7, there were no significant differences within them (6,3and 5) or between them Treatment 1,4,8, and 9, suggesting most treatments (except Treatment 2) perform similarly.

5.	4 pts. The provided function lets you dynamically add a linear model plus one factor from that model and plots a bar chart with letters denoting treatment differences. Use this model to generate the plot shown below. Explain the significance of the letters. 
```{r}
plot_cldbars_onefactor <- function(lm_model, factor) {
  data <- lm_model$model
  variables <- colnames(lm_model$model)
  dependent_var <- variables[1]
  independent_var <- variables[2:length(variables)]

  lsmeans <- emmeans(lm_model, as.formula(paste("~", factor))) # estimate lsmeans 
  Results_lsmeans <- cld(lsmeans, alpha = 0.05, reversed = TRUE, details = TRUE, Letters = letters) # contrast with Tukey adjustment by default.
  
  # Extracting the letters for the bars
  sig.diff.letters <- data.frame(Results_lsmeans$emmeans[,1], 
                                 str_trim(Results_lsmeans$emmeans[,7]))
  colnames(sig.diff.letters) <- c(factor, "Letters")
  
  # for plotting with letters from significance test
  ave_stand2 <- lm_model$model %>%
    group_by(!!sym(factor)) %>%
    dplyr::summarize(
      ave.emerge = mean(.data[[dependent_var]], na.rm = TRUE),
      se = sd(.data[[dependent_var]]) / sqrt(n())
    ) %>%
    left_join(sig.diff.letters, by = factor) %>%
    mutate(letter_position = ave.emerge + 10 * se)
  
  plot <- ggplot(data, aes(x = !! sym(factor), y = !! sym(dependent_var))) + 
    stat_summary(fun = mean, geom = "bar") +
    stat_summary(fun.data = mean_se, geom = "errorbar", width = 0.5) +
    ylab("Number of emerged plants") + 
    geom_jitter(width = 0.02, alpha = 0.5) +
    geom_text(data = ave_stand2, aes(label = Letters, y = letter_position), size = 5) +
    xlab(as.character(factor)) +
    theme_classic()
  
  return(plot)
}

plot_cldbars_onefactor(result2,"Treatment")
```

The letters above each bar represent statistical groupings from a Tukey HSD test (α = 0.05). Treatments sharing the same letter (e.g., "a", "ab") are not significantly different in emergence. For example, Here, the Treatment 2 is labeled "c" (lowest bar), and it is significantly worse than all others. Treatment 7, the second lowest, (labelled "b") is statistically similar to treatments 1, 4, 8, and 9 that are labelled with overlapping letters ( ab ). These treatment labels also overlap with "a" indicating no significant differences with higher treatments ( 3, 5, 6 labeled "a").

6.	2 pts. Generate the gfm .md file along with a .html, .docx, or .pdf. Commit, and push the .md file to github and turn in the .html, .docx, or .pdf to Canvas. Provide me a link here to your github.

[Link to my github]()
  